## **R Libraries: ggplot2**

The R library `ggplot2` is a super useful tool we can use to create visualizations and explore our data.

```{bash}
# install.packages("tidyverse")
# library("tidyverse")
```

More things we can do to get more information on R and what we're doing:

```{bash}
help()
```

We'll need a few more packages for our exploration usage, too. We can install them all at once by using the `c()` function, which creates a list. We'll open these packages with the `library()` function later.

```{bash}
install.packages(c("nycflights13", "gapminder", "Lahman"))
```

Built in to our tidyverse are several data sets that we can play around with. In the below code block, we create a variable `myplot` and save the `mpg` database to it.

```{r}
myplot <- mpg
myplot
```

**Something you should be aware of and have open (perhaps on a second monitor, or you can print them out) is the ggplot2 cheat sheet!** <https://github.com/rstudio/cheatsheets/blob/main/data-visualization-2.1.pdf>

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x=displ, y = hwy, color = class))
```

```{r}
ggplot(data = mpg) + geom_point(mapping = aes(x=displ, y = hwy, color = displ < 5))
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = rgb(0.8, 1, 0.3))) +
  facet_wrap(~ class, nrow = 2)
```

```{r}
#ggplot(data = mpg) +
#  geom_point(mapping = aes(x = displ, y = hwy, shape = class)) +
#  facet_wrap(~ class, nrow = 2)
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = class)) +
  facet_wrap(~ class, nrow = 2)
```

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x= displ, y = hwy, color = drv)) +
  facet_grid(drv ~ .)
```

```{r}
ggplot(data = mpg) +
  geom_smooth(mapping = aes(x= displ, y = hwy, linetype = drv)) +
  geom_point(mapping = aes(x = displ, y = hwy, color = drv))


```

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, linetype = drv)) +
  geom_point(mapping = aes(color = class)) +
  geom_smooth(se = FALSE)
```

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut))
```

Whoa, where'd count come from? **Count is not a variable or attribute of the variable diamonds!** Count comes automatically with a lot of graphs as an attribute. You can verify what I'm saying with `?diamonds` in the console. It shows up because this is a discrete data set, rather than the continuous data set that we have been using.

Any attribute of a data set that is algorithmically calculated is called a *stat*, which is short for a statistical transformation. Many of the **geom** functions have stats built in, and many stats display geoms. For example, the above code block used a geom, but this one uses a stat, and results in an identical chart:

```{r}
ggplot(data = diamonds) +
  stat_count(mapping = aes(x = cut))
# Same as above!!
```

Let's add color here!

```{r}
ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))
```

What if we add to a different variable than cut?

```{r}
pie = ggplot(data = diamonds) +
  geom_bar(mapping = aes(x = cut, fill = clarity),
           position = "fill" # try "dodge" instead of "fill"...graph makes no sense.
  ) +
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)

pie + coord_polar()
```

**For your consideration...** Is the above a good graph?

```{r}
# alt- becomes <- (option- on mac)
```

## Homework

```{r}
# Find a builtin data set for students to explore; use their cheat sheet from stats to label continuous and discrete data; Analyze that data set, create graphs, find a hypothesis
# use graphs to tell me about 3 data sets to tell something Bradon didn't know

View(mtcars)

# Higher hp cars get lower mpg (Duh). Higher number of gears seems to give higher MPG
# I don't like that this gives gradient color for gears, or understand why.
ggplot(data = mtcars, mapping = aes(x = hp, y = mpg)) +
  geom_point(mapping = aes(color = gear)) 

# 4 gears seems to give the highest mpg, but with a wide spread.
ggplot(data = mtcars, mapping = aes(x = gear, y = mpg)) +
  geom_point(mapping = aes(color = hp)) 

# Lower HP gives longer 1/4 mile time (duh)
ggplot(data = mtcars, mapping = aes(x = hp, y = qsec)) +
  geom_point(mapping = aes(color = gear)) +
  geom_smooth(mapping = aes(x = hp, y = qsec))
```

```{r}
# View(morley_mine)
# Why is fill giving a continuous spectrum rather than distinct colors?
# Must be because of third column, Run.
# Nope, even with Run removed it's still like that.
# Maybe because Expt is a number? I could re-label all to Exp 1, Exp 2, etc.
ggplot(data = morley_mine, mapping = aes(x = Expt, y = Speed)) +
  geom_boxplot(mapping = aes(x = Expt, group = Expt, fill = Expt)) +
  labs(x = "Experiment #", y = "Speed of light (km/sec)", title = "Speed of light (km/sec)", subtitle = "Dataset shows c ~= 299,850 km/sec.  Newer data gives c = 299,792.458 km/s")
```

```{r}
View(chickwts)
ggplot(data = chickwts) +
  geom_boxplot(mapping = aes(x = feed, y = weight, fill = feed)) +
  labs(title = 'Chick weight as vs. feed supliment', y = 'weight (g)')
```

```{r}
#View(gmsl)
ggplot(data = GMSL)+
  geom_point(mapping = aes(x = Date, y = sea_lvl_mm)) +
  geom_smooth(mapping = aes(x = Date, y = sea_lvl_mm)) +
  labs(title = "Global Mean Sea Level 1992 - 2013", subtitle = "data courtesy of https://sealevel.colorado.edu/", y = 'Sea level (mm) relative to 1996')
  
```
